---
title: "1.1_scRNASeq_Visualizations"
author: "Matthew Luchs"
date: "2025-07-08"
output: html_document
---

***============ scRNASeq Analysis =============***

**STEP 1: Load Packages, Data, Clean Data**

```{r}
#Install packages
if (!require("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
if (!require("tidyverse", quietly = TRUE))
  install.packages("tidyverse")
if (!require("colorspace", quietly = TRUE))
  install.packages("colorspace")

BiocManager::install("SoupX")

# Install annotation packages for RefSeq mapping
tryCatch({
  BiocManager::install("biomaRt")
  message("biomaRt installed successfully")
}, error = function(e) {
  message("Failed to install biomaRt: ", e$message)
})

tryCatch({
  BiocManager::install("org.Hs.eg.db")
  message("org.Hs.eg.db installed successfully")
}, error = function(e) {
  message("Failed to install org.Hs.eg.db: ", e$message)
})

# Load required libraries
library(Seurat)
library(tidyverse)
library(RColorBrewer)
library(colorspace)
library(ggplot2)

# 1. Load 10X data and perform SoupX background correction
c1_auto_out <- Read10X("/data/pepino/snRNA/C1/outs/filtered_feature_bc_matrix/")
c2_auto_out <- Read10X("/data/pepino/snRNA/C2/outs/filtered_feature_bc_matrix/")

r1_auto_out <- Read10X("/data/pepino/snRNA/R1/outs/filtered_feature_bc_matrix/")
r2_auto_out <- Read10X("/data/pepino/snRNA/R2/outs/filtered_feature_bc_matrix/")

# 2. Create Seurat objects
c1_seurat.obj <- CreateSeuratObject(c1_auto_out, project = 'C1')
c2_seurat.obj <- CreateSeuratObject(c2_auto_out, project = 'C2')
r1_seurat.obj <- CreateSeuratObject(r1_auto_out, project = 'R1')
r2_seurat.obj <- CreateSeuratObject(r2_auto_out, project = 'R2')

# 3. Merge all datasets
all.combined <- merge(c1_seurat.obj, y = c(c2_seurat.obj, r1_seurat.obj, r2_seurat.obj),
          add.cell.ids = c("C1", "C2", "R1", "R2"), project = "ALL")

# 4. Add sample metadata
all.combined$log10GenesPerUMI <- log10(all.combined$nFeature_RNA) / log10(all.combined$nCount_RNA)

all.combined_metadata <- all.combined@meta.data
all.combined_metadata$cells <- rownames(all.combined_metadata)
all.combined_metadata$sample <- NA
all.combined_metadata$sample[grep("^C1_", rownames(all.combined_metadata))] <- "C1"
all.combined_metadata$sample[grep("^C2_", rownames(all.combined_metadata))] <- "C2"
all.combined_metadata$sample[grep("^R1_", rownames(all.combined_metadata))] <- "R1"
all.combined_metadata$sample[grep("^R2_", rownames(all.combined_metadata))] <- "R2"

all.combined@meta.data <- all.combined_metadata

# 5. Filter, normalize, and find variable features
all.combined <- subset(all.combined, subset = nFeature_RNA > 50 & nFeature_RNA < 7500 & nCount_RNA < 40000)
all.combined <- NormalizeData(all.combined)
all.combined <- FindVariableFeatures(all.combined, selection.method = "vst", nfeatures = 2000)
```

**Step 2: PCA/UMAP**

```{r}
# 6. Scaling, PCA, neighbors, clustering, UMAP
all.combined <- ScaleData(all.combined)
all.combined <- RunPCA(all.combined)
ElbowPlot(all.combined)
all.combined <- FindNeighbors(all.combined, dims = 1:20)
all.combined <- FindClusters(all.combined, resolution = 1.84)
all.combined <- RunUMAP(all.combined, dims = 1:20)
```

**Step 3: Visualizaiton**

```{r}
# 8a. UMAP for all clusters from all samples in the same plot
DimPlot(all.combined, reduction = "umap", group.by = "seurat_clusters", label = TRUE, label.size = 4) +
  ggtitle("UMAP: All Clusters from All Samples") +
  theme(plot.title = element_text(hjust = 0.5))

# 8b. Create subsets for C1/C2 and R1/R2
c1_c2_subset <- subset(all.combined, subset = sample %in% c("C1", "C2"))
r1_r2_subset <- subset(all.combined, subset = sample %in% c("R1", "R2"))

# Determine which sample has fewer cells in each subset
c1_cells <- sum(c1_c2_subset$sample == "C1")
c2_cells <- sum(c1_c2_subset$sample == "C2")
r1_cells <- sum(r1_r2_subset$sample == "R1")
r2_cells <- sum(r1_r2_subset$sample == "R2")

print(paste("C1 cells:", c1_cells, "C2 cells:", c2_cells))
print(paste("R1 cells:", r1_cells, "R2 cells:", r2_cells))

# Create the smaller sample objects
c1_c2_smaller <- ifelse(c1_cells < c2_cells, "C1", "C2")
r1_r2_smaller <- ifelse(r1_cells < r2_cells, "R1", "R2")


#UMAP Plots by Group (C or R)

# Your original plots with labels added
p1 <- DimPlot(c1_c2_subset, reduction = "umap", group.by = "sample",
              order = c1_c2_smaller) +  # This puts the smaller sample on top
  ggtitle("UMAP: C1/C2 Samples") +
  theme(plot.title = element_text(hjust = 0.5))

# Add cluster labels
embeddings <- Embeddings(c1_c2_subset, reduction = "umap")
clusters <- c1_c2_subset$seurat_clusters
centroids <- do.call(rbind, lapply(split(as.data.frame(embeddings), clusters), function(x) apply(x, 2, mean)))
p1 <- p1 + annotate("text", x = centroids[,1], y = centroids[,2], 
                    label = rownames(centroids), size = 4, color = "black")
print(p1)

# Just Regenerating
p2 <- DimPlot(r1_r2_subset, reduction = "umap", group.by = "sample",
              order = r1_r2_smaller) +  # This puts the smaller sample on top
  ggtitle("UMAP: R1/R2 Samples") +
  theme(plot.title = element_text(hjust = 0.5))

# Add cluster labels
embeddings2 <- Embeddings(r1_r2_subset, reduction = "umap")
clusters2 <- r1_r2_subset$seurat_clusters
centroids2 <- do.call(rbind, lapply(split(as.data.frame(embeddings2), clusters2), function(x) apply(x, 2, mean)))
p2 <- p2 + annotate("text", x = centroids2[,1], y = centroids2[,2], 
                    label = rownames(centroids2), size = 4, color = "black")
print(p2)

# 8c. Representative cluster distribution plots
# Create table of cluster vs sample
table_data <- table(all.combined@meta.data$seurat_clusters, all.combined@meta.data$sample)

# PLOT 1: Proportion of each SAMPLE found in each cluster (more representative)
# Convert to proportions within each sample (columns sum to 1)
prop_by_sample <- prop.table(table_data, margin = 2) * 100

# Convert to dataframe for ggplot
df_by_sample <- as.data.frame(as.table(prop_by_sample))
colnames(df_by_sample) <- c("Cluster", "Sample", "Percentage")

# Define custom colors: C1=green, C2=blue, R1=red, R2=purple
sample_colors <- c("C1" = "#2E8B57", "C2" = "#4169E1", "R1" = "#DC143C", "R2" = "#9932CC")

# Create representative barplot showing what % of each sample is in each cluster
ggplot(df_by_sample, aes(x = Cluster, y = Percentage, fill = Sample)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Percent of Each Sample in Each Cluster", 
       x = "Cluster", 
       y = "% of Sample's Total Cells",
       fill = "Sample",
       subtitle = "Shows what proportion of each sample's cells are found in each cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, max(df_by_sample$Percentage) * 1.1))

# Convert to proportions within each cluster (rows sum to 1)
prop_by_cluster <- prop.table(table_data, margin = 1) * 100

# Convert to dataframe for ggplot
df_stacked <- as.data.frame(as.table(prop_by_cluster))
colnames(df_stacked) <- c("Cluster", "Sample", "Percentage")

# Create stacked barplot
ggplot(df_stacked, aes(x = Cluster, y = Percentage, fill = Sample)) +
  geom_bar(stat = "identity", position = "stack") +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Sample Composition by Cluster (Raw Proportions)", 
       x = "Cluster", 
       y = "Percentage of Cells in Cluster",
       fill = "Sample",
       subtitle = "Shows what % of each cluster is made up of each sample (may be biased by sample size)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_cartesian(ylim = c(0, 100))
```

**Step 4: Subcluster Visuals - Muscle Group (Clusters 3,10,21)**

```{r}
# Extract cells from clusters 3, 10, and 21
target_clusters <- c("3", "10", "21")
clusters_subset <- subset(all.combined, subset = seurat_clusters %in% target_clusters)
print(paste("Total cells in clusters 3, 10, 21:", ncol(clusters_subset)))

# Save the original cluster information before it gets overwritten
clusters_subset$original_cluster <- clusters_subset$seurat_clusters

# Re-scale, run PCA, and find new clusters/UMAP
clusters_subset <- ScaleData(clusters_subset)
clusters_subset <- RunPCA(clusters_subset, features = VariableFeatures(all.combined))
clusters_subset <- FindNeighbors(clusters_subset, dims = 1:15)
clusters_subset <- FindClusters(clusters_subset, resolution = 0.8)
clusters_subset <- RunUMAP(clusters_subset, dims = 1:15)
```

**Step 4B - Muscle UMAP**

```{r}
# 1. UMAP colored by new subclusters
DimPlot(clusters_subset, reduction = "umap", label = TRUE, label.size = 4) +
  ggtitle("Muscle UMAP: Colored by New Subclusters") +
  theme(plot.title = element_text(hjust = 0.5))

# 2. UMAP colored by sample (C1, C2, R1, R2)
DimPlot(clusters_subset, reduction = "umap", group.by = "sample") +
  ggtitle("Muscle UMAP: Colored by Sample") +
  theme(plot.title = element_text(hjust = 0.5))

# 3. UMAP colored by original cluster (3, 10, or 21)
DimPlot(clusters_subset, reduction = "umap", group.by = "original_cluster") +
  ggtitle("Muscle UMAP: Colored by Original Cluster (3, 10, 21)") +
  theme(plot.title = element_text(hjust = 0.5))
```

**Step 4C - Muscle Propotional Visuals**

```{r}
# Create data for sample composition bar plot
# Calculate percentage of each sample in each new cluster
sample_cluster_counts <- table(clusters_subset$seurat_clusters, clusters_subset$sample)
sample_totals <- colSums(sample_cluster_counts)

# Convert to percentages (what % of each sample is in each cluster)
df_by_sample <- data.frame()
for(sample in names(sample_totals)) {
  for(cluster in rownames(sample_cluster_counts)) {
    count <- sample_cluster_counts[cluster, sample]
    percentage <- (count / sample_totals[sample]) * 100
    df_by_sample <- rbind(df_by_sample, data.frame(
      Cluster = cluster,
      Sample = sample,
      Count = count,
      Percentage = percentage
    ))
  }
}

# Create data for original cluster composition bar plot
# Calculate percentage of each original cluster in each new cluster
original_cluster_counts <- table(clusters_subset$seurat_clusters, clusters_subset$original_cluster)
original_totals <- colSums(original_cluster_counts)

# Convert to percentages (what % of each original cluster is in each new cluster)
df_by_original <- data.frame()
for(orig_cluster in names(original_totals)) {
  for(new_cluster in rownames(original_cluster_counts)) {
    count <- original_cluster_counts[new_cluster, orig_cluster]
    percentage <- (count / original_totals[orig_cluster]) * 100
    df_by_original <- rbind(df_by_original, data.frame(
      New_Cluster = new_cluster,
      Original_Cluster = paste("Cluster", orig_cluster),
      Count = count,
      Percentage = percentage
    ))
  }
}

# Define colors for samples (adjust as needed)
sample_colors <- c("C1" = "#FF6B6B", "C2" = "#4ECDC4", "R1" = "#45B7D1", "R2" = "#96CEB4")

# Define colors for original clusters
original_colors <- c("Cluster 3" = "#FF9999", "Cluster 10" = "#66B2FF", "Cluster 21" = "#99FF99")

# 4. Bar plot showing what % of each sample is in each new cluster
ggplot(df_by_sample, aes(x = Cluster, y = Percentage, fill = Sample)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Percent of Each Sample in Each New Cluster", 
       x = "New Cluster", 
       y = "% of Sample's Total Cells",
       fill = "Sample   ",
       subtitle = "Shows what proportion of each sample's cells are found in each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, max(df_by_sample$Percentage) * 1.1))

# 5. Bar plot showing what % of each original cluster is in each new cluster
ggplot(df_by_original, aes(x = New_Cluster, y = Percentage, fill = Original_Cluster)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_fill_manual(values = original_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Percent of Each Original Cluster in Each New Cluster", 
       x = "New Cluster", 
       y = "% of Original Cluster's Total Cells",
       fill = "Original Cluster   ",
       subtitle = "Shows what proportion of each original cluster's cells are found in each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, max(df_by_original$Percentage) * 1.1))

# Create comprehensive data frame with all combinations
cluster_composition <- data.frame(
  new_cluster = clusters_subset$seurat_clusters,
  original_cluster = clusters_subset$original_cluster,
  sample = clusters_subset$sample
)

# Calculate counts for each combination
composition_counts <- cluster_composition %>%
  group_by(new_cluster, original_cluster, sample) %>%
  summarise(count = n(), .groups = 'drop')

# Define colors
sample_colors <- c("C1" = "firebrick", "C2" = "orange", "R1" = "deepskyblue", "R2" = "purple")
original_colors <- c("3" = "firebrick", "10" = "blue", "21" = "purple")

# Option 1: Stacked by sample (shows sample composition within each new cluster)
# Calculate percentages within each new cluster
composition_by_sample <- composition_counts %>%
  group_by(new_cluster) %>%
  mutate(total_in_cluster = sum(count),
         percentage = (count / total_in_cluster) * 100)

ggplot(composition_by_sample, aes(x = new_cluster, y = percentage, fill = sample)) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Sample Composition Within Each New Cluster", 
       x = "New Cluster", 
       y = "Percentage of Cells in Cluster",
       fill = "Sample",
       subtitle = "Shows what samples make up each new cluster (stacked to 100%)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))

# Option 2: Stacked by original cluster (shows original cluster composition within each new cluster)
# Calculate percentages within each new cluster
composition_by_original <- composition_counts %>%
  group_by(new_cluster) %>%
  mutate(total_in_cluster = sum(count),
         percentage = (count / total_in_cluster) * 100)

ggplot(composition_by_original, aes(x = new_cluster, y = percentage, fill = original_cluster)) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) +
  scale_fill_manual(values = original_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Original Cluster Composition Within Each New Cluster", 
       x = "New Cluster", 
       y = "Percentage of Cells in Cluster",
       fill = "Original Cluster",
       subtitle = "Shows what original clusters make up each new cluster (stacked to 100%)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))
```



**Step 5: Subcluster Visuals - Nervous System**

```{r}
# Extract cells from clusters 0,5,19,22,23,26,27,32,33
target_clusters <- c("0","5","19","22","23","26","27","32","33")
clusters_subset <- subset(all.combined, subset = seurat_clusters %in% target_clusters)
print(paste("Total cells in clusters 0,5,19,22,23,26,27,32,33:", ncol(clusters_subset)))

# Save the original cluster information before it gets overwritten
clusters_subset$original_cluster <- clusters_subset$seurat_clusters

# Re-scale, run PCA, and find new clusters/UMAP
clusters_subset <- ScaleData(clusters_subset)
clusters_subset <- RunPCA(clusters_subset, features = VariableFeatures(all.combined))
clusters_subset <- FindNeighbors(clusters_subset, dims = 1:15)
clusters_subset <- FindClusters(clusters_subset, resolution = 0.8)
clusters_subset <- RunUMAP(clusters_subset, dims = 1:15)
```

**Step 5B - Nervous System UMAP**

```{r}
# 1. UMAP colored by new subclusters
DimPlot(clusters_subset, reduction = "umap", label = TRUE, label.size = 4) +
  ggtitle("NS UMAP: Colored by New Subclusters") +
  theme(plot.title = element_text(hjust = 0.5))

# 2. UMAP colored by sample (C1, C2, R1, R2)
DimPlot(clusters_subset, reduction = "umap", group.by = "sample") +
  ggtitle("NS UMAP: Colored by Sample") +
  theme(plot.title = element_text(hjust = 0.5))

# 3. UMAP colored by original cluster (0,5,19,22,23,26,27,32,33)
DimPlot(clusters_subset, reduction = "umap", group.by = "original_cluster") +
  ggtitle("NS UMAP: Colored by Original Cluster") +
  theme(plot.title = element_text(hjust = 0.5))
```

**Step 5C - Nervous System Proportional Visuals**

```{r}
# Create data for sample composition bar plot
# Calculate percentage of each sample in each new cluster
sample_cluster_counts <- table(clusters_subset$seurat_clusters, clusters_subset$sample)
sample_totals <- colSums(sample_cluster_counts)

# Convert to percentages (what % of each sample is in each cluster)
df_by_sample <- data.frame()
for(sample in names(sample_totals)) {
  for(cluster in rownames(sample_cluster_counts)) {
    count <- sample_cluster_counts[cluster, sample]
    percentage <- (count / sample_totals[sample]) * 100
    df_by_sample <- rbind(df_by_sample, data.frame(
      Cluster = cluster,
      Sample = sample,
      Count = count,
      Percentage = percentage
    ))
  }
}

# Create data for original cluster composition bar plot
# Calculate percentage of each original cluster in each new cluster
original_cluster_counts <- table(clusters_subset$seurat_clusters, clusters_subset$original_cluster)
original_totals <- colSums(original_cluster_counts)

# Convert to percentages (what % of each original cluster is in each new cluster)
df_by_original <- data.frame()
for(orig_cluster in names(original_totals)) {
  for(new_cluster in rownames(original_cluster_counts)) {
    count <- original_cluster_counts[new_cluster, orig_cluster]
    percentage <- (count / original_totals[orig_cluster]) * 100
    df_by_original <- rbind(df_by_original, data.frame(
      New_Cluster = new_cluster,
      Original_Cluster = paste("Cluster", orig_cluster),
      Count = count,
      Percentage = percentage
    ))
  }
}

# Define colors for samples (adjust as needed)
sample_colors <- c("C1" = "#FF6B6B", "C2" = "#4ECDC4", "R1" = "#45B7D1", "R2" = "#96CEB4")

# Define colors for original clusters
original_colors <- c("Cluster 0" = "red","Cluster 5" = "blue","Cluster 19" = "green","Cluster 22" = "purple","Cluster 23" = "orange","Cluster 26" = "yellow","Cluster 27" = "brown","Cluster 32" = "pink","Cluster 33" = "darkgray")

# 4. Bar plot showing what % of each sample is in each new cluster
ggplot(df_by_sample, aes(x = reorder(Cluster, as.numeric(Cluster)), y = Percentage, fill = Sample)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Percent of Each Sample in Each New Cluster", 
       x = "New Cluster", 
       y = "% of Sample's Total Cells",
       fill = "Sample   ",
       subtitle = "Shows what proportion of each sample's cells are found in each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, max(df_by_sample$Percentage) * 1.1))

# 5. Bar plot showing what % of each original cluster is in each new cluster
ggplot(df_by_original, aes(x = reorder(New_Cluster, as.numeric(New_Cluster)), y = Percentage, fill = Original_Cluster)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_fill_manual(values = original_colors) +
  theme_minimal(base_size = 9) +
  labs(title = "Percent of Each Original Cluster in Each New Cluster", 
       x = "New Cluster", 
       y = "% of Original Cluster Total Cells",
       fill = "Original Cluster   ",
       subtitle = "Proportion of each original cluster's cells in each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, max(df_by_original$Percentage) * 1.1))

# Create comprehensive data frame with all combinations
cluster_composition <- data.frame(
  new_cluster = clusters_subset$seurat_clusters,
  original_cluster = clusters_subset$original_cluster,
  sample = clusters_subset$sample
)

# Calculate counts for each combination
composition_counts <- cluster_composition %>%
  group_by(new_cluster, original_cluster, sample) %>%
  summarise(count = n(), .groups = 'drop')

# Option 1: Stacked by sample (shows sample composition within each new cluster)
# Calculate percentages within each new cluster
composition_by_sample <- composition_counts %>%
  group_by(new_cluster) %>%
  mutate(total_in_cluster = sum(count),
         percentage = (count / total_in_cluster) * 100)

ggplot(composition_by_sample, aes(x = new_cluster, y = percentage, fill = sample)) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Sample Composition Within Each New Cluster", 
       x = "New Cluster", 
       y = "Percentage of Cells in Cluster",
       fill = "Sample",
       subtitle = "Shows what samples make up each new cluster (stacked to 100%)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))

# Option 2: Stacked by original cluster (shows original cluster composition within each new cluster)
# Calculate percentages within each new cluster
composition_by_original <- composition_counts %>%
  group_by(new_cluster) %>%
  mutate(total_in_cluster = sum(count),
         percentage = (count / total_in_cluster) * 100)

ggplot(composition_by_original, aes(x = new_cluster, y = percentage, fill = original_cluster)) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) +
  #scale_fill_manual(values = original_colors) + ##Colors not wokring, letting R choose 
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 12) +
  labs(title = "Original Cluster Composition Within New Clusters", 
       x = "New Cluster", 
       y = "Percentage of Cells in Cluster",
       fill = "Original Cluster",
       subtitle = "Shows what original clusters make up each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))
```

**Step 6: Subcluster Visuals - Connective Tissue**

```{r}
# Extract cells from clusters 7,11,15,25,29
target_clusters <- c("7","11","15","25","29")
clusters_subset <- subset(all.combined, subset = seurat_clusters %in% target_clusters)
print(paste("Total cells in clusters 7,11,15,25,29:", ncol(clusters_subset)))

# Save the original cluster information before it gets overwritten
clusters_subset$original_cluster <- clusters_subset$seurat_clusters

# Re-scale, run PCA, and find new clusters/UMAP
clusters_subset <- ScaleData(clusters_subset)
clusters_subset <- RunPCA(clusters_subset, features = VariableFeatures(all.combined))
clusters_subset <- FindNeighbors(clusters_subset, dims = 1:15)
clusters_subset <- FindClusters(clusters_subset, resolution = 0.8)
clusters_subset <- RunUMAP(clusters_subset, dims = 1:15)
```

**Step 6B - Connective Tissue UMAP**

```{r}
# 1. UMAP colored by new subclusters
DimPlot(clusters_subset, reduction = "umap", label = TRUE, label.size = 4) +
  ggtitle("ConTis UMAP: Colored by New Subclusters") +
  theme(plot.title = element_text(hjust = 0.5))

# 2. UMAP colored by sample (C1, C2, R1, R2)
DimPlot(clusters_subset, reduction = "umap", group.by = "sample") +
  ggtitle("ConTis UMAP: Colored by Sample") +
  theme(plot.title = element_text(hjust = 0.5))

# 3. UMAP colored by original cluster (7,11,15,25,29)
DimPlot(clusters_subset, reduction = "umap", group.by = "original_cluster") +
  ggtitle("ConTis UMAP: Colored by Original Cluster") +
  theme(plot.title = element_text(hjust = 0.5))
```

**Step 6C - Connective Proportional Visuals**

```{r}
# Create data for sample composition bar plot
# Calculate percentage of each sample in each new cluster
sample_cluster_counts <- table(clusters_subset$seurat_clusters, clusters_subset$sample)
sample_totals <- colSums(sample_cluster_counts)

# Convert to percentages (what % of each sample is in each cluster)
df_by_sample <- data.frame()
for(sample in names(sample_totals)) {
  for(cluster in rownames(sample_cluster_counts)) {
    count <- sample_cluster_counts[cluster, sample]
    percentage <- (count / sample_totals[sample]) * 100
    df_by_sample <- rbind(df_by_sample, data.frame(
      Cluster = cluster,
      Sample = sample,
      Count = count,
      Percentage = percentage
    ))
  }
}

# Create data for original cluster composition bar plot
# Calculate percentage of each original cluster in each new cluster
original_cluster_counts <- table(clusters_subset$seurat_clusters, clusters_subset$original_cluster)
original_totals <- colSums(original_cluster_counts)

# Convert to percentages (what % of each original cluster is in each new cluster)
df_by_original <- data.frame()
for(orig_cluster in names(original_totals)) {
  for(new_cluster in rownames(original_cluster_counts)) {
    count <- original_cluster_counts[new_cluster, orig_cluster]
    percentage <- (count / original_totals[orig_cluster]) * 100
    df_by_original <- rbind(df_by_original, data.frame(
      New_Cluster = new_cluster,
      Original_Cluster = paste("Cluster", orig_cluster),
      Count = count,
      Percentage = percentage
    ))
  }
}

# Define colors for samples (adjust as needed)
sample_colors <- c("C1" = "#FF6B6B", "C2" = "#4ECDC4", "R1" = "#45B7D1", "R2" = "#96CEB4")

# Define colors for original clusters
original_colors <- c("Cluster 0" = "red","Cluster 5" = "blue","Cluster 19" = "green","Cluster 22" = "purple","Cluster 23" = "orange","Cluster 26" = "yellow","Cluster 27" = "brown","Cluster 32" = "pink","Cluster 33" = "darkgray")

# 4. Bar plot showing what % of each sample is in each new cluster
ggplot(df_by_sample, aes(x = reorder(Cluster, as.numeric(Cluster)), y = Percentage, fill = Sample)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Percent of Each Sample in Each New Cluster", 
       x = "New Cluster", 
       y = "% of Sample's Total Cells",
       fill = "Sample   ",
       subtitle = "Shows what proportion of each sample's cells are found in each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, max(df_by_sample$Percentage) * 1.1))

# 5. Bar plot showing what % of each original cluster is in each new cluster
ggplot(df_by_original, aes(x = reorder(New_Cluster, as.numeric(New_Cluster)), y = Percentage, fill = Original_Cluster)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.8) +
  scale_fill_manual(values = original_colors) +
  theme_minimal(base_size = 9) +
  labs(title = "Percent of Each Original Cluster in Each New Cluster", 
       x = "New Cluster", 
       y = "% of Original Cluster Total Cells",
       fill = "Original Cluster   ",
       subtitle = "Proportion of each original cluster's cells in each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, max(df_by_original$Percentage) * 1.1))

# Create comprehensive data frame with all combinations
cluster_composition <- data.frame(
  new_cluster = clusters_subset$seurat_clusters,
  original_cluster = clusters_subset$original_cluster,
  sample = clusters_subset$sample
)

# Calculate counts for each combination
composition_counts <- cluster_composition %>%
  group_by(new_cluster, original_cluster, sample) %>%
  summarise(count = n(), .groups = 'drop')

# Option 1: Stacked by sample (shows sample composition within each new cluster)
# Calculate percentages within each new cluster
composition_by_sample <- composition_counts %>%
  group_by(new_cluster) %>%
  mutate(total_in_cluster = sum(count),
         percentage = (count / total_in_cluster) * 100)

ggplot(composition_by_sample, aes(x = new_cluster, y = percentage, fill = sample)) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) +
  scale_fill_manual(values = sample_colors) +
  theme_minimal(base_size = 12) +
  labs(title = "Sample Composition Within Each New Cluster", 
       x = "New Cluster", 
       y = "Percentage of Cells in Cluster",
       fill = "Sample",
       subtitle = "Shows what samples make up each new cluster (stacked to 100%)") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))

# Option 2: Stacked by original cluster (shows original cluster composition within each new cluster)
# Calculate percentages within each new cluster
composition_by_original <- composition_counts %>%
  group_by(new_cluster) %>%
  mutate(total_in_cluster = sum(count),
         percentage = (count / total_in_cluster) * 100)

ggplot(composition_by_original, aes(x = new_cluster, y = percentage, fill = original_cluster)) +
  geom_bar(stat = "identity", position = "stack", width = 0.8) +
  #scale_fill_manual(values = original_colors) + ##Colors not wokring, letting R choose 
  scale_fill_brewer(palette = "Set1") +
  theme_minimal(base_size = 12) +
  labs(title = "Original Cluster Composition Within New Clusters", 
       x = "New Cluster", 
       y = "Percentage of Cells in Cluster",
       fill = "Original Cluster",
       subtitle = "Shows what original clusters make up each new cluster") +
  theme(plot.title = element_text(hjust = 0.5),
        plot.subtitle = element_text(hjust = 0.5, size = 10),
        axis.text.x = element_text(angle = 0, hjust = 0.5),
        panel.grid.major.x = element_line(color = "gray90", size = 0.5),
        panel.grid.minor = element_blank(),
        legend.position = "bottom") +
  coord_cartesian(ylim = c(0, 100))
```
